\documentclass{sig-alternate-05-2015}
  \pdfpagewidth=8.5truein
  \pdfpageheight=11truein
  
  \usepackage{graphicx}
  \usepackage[portuguese, english]{babel}
  \graphicspath{ {images/} }
  \usepackage[utf8]{inputenc}
  \usepackage[justification=centering]{caption}
  \usepackage{amsmath}
  \usepackage{multirow}
  \usepackage{makecell}
  \usepackage{cite}
  \usepackage{booktabs}
  %\usepackage{changes}
  \usepackage[final]{changes}
  \usepackage{enumerate}
%  \usepackage{subfigure}
\usepackage{caption}
\usepackage{subcaption}

\begin{document}

\hyphenation{semi-au-to-mat-ic}

% Copyright
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\doi{http://dx.doi.org/xx.xxxx/xxxxxxx.xxxxxxx}

% ISBN
\isbn{978-1-4503-3739-7/16/04}

%Conference
%\conferenceinfo{PLDI '13}{June 16--19, 2013, Seattle, WA, USA}

\acmPrice{\$15.00}

%
% --- Author Metadata here ---
\conferenceinfo{SAC'16,}{ April 4-8, 2016, Pisa, Italy}
\CopyrightYear{2016} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Exploring the Combination of Software Visualization and Data Clustering in the Software Architecture Recovery Process}


\numberofauthors{1} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
%\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%ah
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
%\alignauthor
%Renato Paiva, Genaína N. Rodrigues,  Rodrigo Bonif\'{a}cio\\
%      \affaddr{Computer Science Department}\\
%     \affaddr{University of Brasília}\\
%    \affaddr{Brasilia, Brazil}\\
%    \email{\{renatoedesio}@unb.br\},\\
%   \{genaina, rbonifacio\}@cic.unb.br}



\maketitle
\begin{abstract}
Modernizing a legacy system is a costly process that requires deep understanding of the system architecture and its components. Without an understanding of the software architecture that will be rewritten, the entire process of reengineering can fail. \deleted{When there is absence of architectural documents, it is important to have a recovery process of architecture that allows the complete understanding of the software. Such process involves mapping of source code entities in high-level models.} \replaced{For this reason, semi-automatic and automatic techniques for architecture recovery have been active focuses of research. However, there are still important improvements that need to be addressed on this field of research w.r.t achieving a more accurate architecture recovery process.}{Previous work using visualization and clustering techniques has been proposed and extensively used. However, their accuracy to reconstruct the architecture alone has shown to be not satisfactory enough.} In this work, we propose to explore if an approach where visualization and clustering applied together can provide a higher accuracy on the software architecture recovery process. An experimental study was conducted in a industrial environment to empirically evaluate our investigation in four commercial systems. Our results indicated a statistically significant increase in the accuracy of the recovered architectures in all cases.
\end{abstract}


%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}


%
% End generated code
%

%
%  Use this command to print the description
%
%\printccsdesc

% We no longer use \terms command
%\terms{Theory}

\keywords{Software Architecture Recovery; Software Visualization; Data Clustering}

\input{introduction}
\input{related}
\input{rationale}
\input{experiment}
\input{conclusion}

\bibliographystyle{abbrv}
\bibliography{bibliografia}  % sigproc.bib is the name of the Bibliography in this case


\end{document}
